\bf Title

### 2024-01-16 article/paper_padic_spec.tex
other modifs

\newenvironment{proof}{\hbox{}\vspace{-0.5cm} {\bf Proof:}}{\hfill $\Box$ \\}

@@

\begin{definition}
  \label{def_polyhedra}
  A \emph{polyhedron} is a subset $P \subset K^n$ of the form
  $$
  P = \{x \in K^n : \val(\ell_1 \oplus \cdots \oplus \ell_d) \geq 0 \text{ and }
  \val(m_1 \oplus \cdots \oplus m_e) = +\infty\},
  $$
  for some $\ell_1,\ldots,\ell_d,m_1,\ldots,m_e \in K[x]_1$, where $\val,\geq$ and $=$ are applied entrywise.
  We always assume without loss of generality that $m_1,\ldots,m_e$ are $K$-linearly independent.
\end{definition}

@@

\begin{definition}[Smith Normal Form]\label{smith_nf}
  Let $M \in \allmat_{m,n}(K)$ be of rank $r$. There exist unique integers $a_1 \leq \cdots \leq a_r$, and
  matrices $P \in \GL_n(\OK)$ and $Q \in \GL_m(\OK)$, such that
  $$
  S := QMP^{-1} = \pi^{a_1} \oplus \pi^{a_2} \oplus \cdots \oplus \pi^{a_r} \oplus {\bf 0}_{\min(m,n)-r}
  $$
  where ${\bf 0}_{\min(m,n)-r}$ is a null square matrix of size $\min(m,n)-r$.
  The matrix $S$ is called the \emph{Smith Normal Form (SNF)} of $M$, and the elements $\pi^{a_i}$ the
  \emph{invariant factors}.
\end{definition}
\begin{remark}
  The valuation of the first invariant factor of the SNF of a matrix
  $M \in \allmat_{n,n} \left( K \right)$ is the minimum of the valuations of the entries of
  $M$. In particular, the SNF of $M \in \allmat_n \left( \OK \right)$ has entries in $\OK$.
\end{remark}

@@

The \emph{associated diagonal linear programming problem} of a linear program is the iteration of the \ref{dLP} problem defined by $b' = Qb$, $c' = \left(P^{-1}\right)^T c$ and where $S$ is the Smith normal form of $A$ and $P \in GL_m\left( \OK \right)$ and $ Q \in GL_n\left( \OK \right)$ are such that $A = Q^{-1} S P$.

@@

\subsection{Definition}
%\newcommand\mat{postive semidefinite matrix } 
\newcommand\Mat{Positive semidefinite matrix }
\newcommand\mats{positive semidefinite matrices }
\newcommand\Mats{positive semidefinite matrices }
\begin{definition}
	We call ... any matrix $M \in \mathcal{M}_n\left( K \right) $ whose eigenvalues have nonegative valuation (in $\overline{K}$) .

	We will denote as $ \mathcal{M}_n^+\left( K \right)$ the set of \mats and we will write $M \succeq 0$ if $M \in \mathcal{M}_n^+\left( K\right) $.
\end{definition}
\begin{theorem}
	\label{caracsdp}
	Characterization of the \mats

	A matrix is positive semidefinite if and only if its characteristic polynomial have coefficients in $\OK$.

\end{theorem}

\begin{corollary}
	$\mathcal{M}_n(\OK) \subset \mathcal{M}_n^+\left( K \right)$ 
\end{corollary}

\begin{proof}
$\OK$ is a ring therefore the characteristic polynomial of matrix whose coefficients are $p$-adic integers have coefficients ins $Pp$. We conclude by \ref{caracsdp} .
\end{proof}

\begin{remark}
	In general, the reverse inclusion is false. For instance the characteristic polynomial of $M = \begin{pmatrix} 5 + \frac{3}{5} & \frac{4}{5} \\ \frac{4}{5} & -\frac{3}{5} \end{pmatrix} $ is $\chi_M = X^2  - 5 X - 4$. However when seen as an element of $\mathbb{Q}_5$, $\chi_M \in \mathbb{Z}_5[X]$ then $M \in \mathcal{M}_2^+\left( \mathbb{Q}_5 \right)$ but no coefficient of $M$ is in $\mathbb{Z}_5 $
\end{remark}

\begin{proposition}
	The set $\mathcal{M}_n^+\left( K \right)$ is :
	\begin{enumerate}%[label = \textit{\roman*}.]
		\item open
		\item closed
		%\item a convex cone
	\end{enumerate}
\end{proposition}

\textit{Preuve :}  
Firstly, $\OK[X]$ is closed and open in $\mathbb{Q}_{ p }[X] $ equiped with the infinitiy norm $\|.\|_\infty : P = \sum_{k=1}^{n} a_k X^k \to \sup |a_k|_p$. $\OK$ is indeed the open unit ball for $\|\cdot \|_\infty$ which defines a discrete distance for which open balls are also closed.
Furthermore, $\mathcal{M}^+_{n}\left(K \right)  = \chi^{-1}( \OK[X]) $ where $\chi$ is the function that maps its characteristic polynomial to a matrix which is continuous.

%Le $iii$. est laissé en exercice au lecteur. 
% Je ne parle plus de convexité dans le rapport mais on ne sait jamais
%\input{convexemple.tex}

\begin{definition}
  We call \emph{spectrahedron} the intersection of $\mathcal{M}_n^+\left( K \right) $ with an affine plan $\mathcal{L}$ of $\mathcal{M}_n\left( K \right) $.
\end{definition}

\begin{remark}
	As in the real case, the spectrahedron associated with the affine plan $\mathcal{L}$ will often be identified with its preimage in $K^s$ (i.e. the set of vectors $x \in K ^s$ so that the linear matrix $A(x) = A_0 + x_1A_1 + \ldots + x_sA_s$ is positive semidefinite with $A_0,A_1,\ldots,A_s$ spanning $\mathcal{L}$). 
\end{remark}

The latter allows gives us the following property :
\begin{proposition}
	A polyhedron is a spectrahedron.
\end{proposition}
\begin{proof}
	Same as in the real case (maybe modify it or add the link to real proof).
\end{proof}

@@

	We call \emph{annulus} a set $A$ defined as $C = \{x \in K  | a\le \val\left(x\right) \le b\} $ for $a < b$ two nonnegative real numbers.
\end{definition}

@@

\textit{Preuve :}
Let $A$ be the annuli defined by $a<b \in R^*_+$.

@@

Let us show that $x \in C$ if and only if  $\exists y \in K \left( x,y \right) \in \mathcal{S}$.

@@

Therefore $x \in C$ if and only if there exists $y \in K $ so that $(x, y)$ verifies \ref{eq:caracsdp}. Si $x, y$ verifies indeed \ref{caracsdp} then $\val\left(x\right)\ge a$, $\val\left(y\right)\ge -b$ and $p^{-2} \left( xy-1 \right) \ge 0$ implies that $\val\left(x\right)+\val\left(y\right)=\val\left(xy\right) = \val\left(-1\right) =0$ and therefore that $\val\left(x\right)=-\val\left(y\right)\le -(-b)  =b$, so $x \in $. Reciprocally if $x \in C$ then $(x,x^{-1}) $ verifies \ref{eq:caracsdp}  .

>>>>

\newenvironment{proof}{\hbox{}\vspace{-0.8cm} {\bf Proof:}}{\hfill $\Box$}

@@

\begin{definition}
  \label{def_polyhedra}
  A \emph{polyhedron} is a subset $P \subset K^n$ of the form
  $$
  P = \{x \in K^n : \val(\ell_1 \oplus \cdots \oplus \ell_d) \geq 0 \text{ and }
  \val(m_1 \oplus \cdots \oplus m_e) = +\infty\},
  $$
  for some $\ell_1,\ldots,\ell_d,m_1,\ldots,m_e \in K[x]_1$, where $\val,\geq$ and $=$ are applied entrywise.
  We always assume without loss of generality that $m_1,\ldots,m_e$ are affinely independent.
\end{definition}

@@

\begin{definition}[Smith Normal Form]\label{smith_nf}
  Let $M \in \allmat_{m,n}(K)$ be of rank $r$. There exist unique integers $a_1 \leq \cdots \leq a_r$, and
  matrices $P \in \GL_n(\OK)$ and $Q \in \GL_m(\OK)$, such that
  $$
  S := QMP^{-1} = \pi^{a_1} \oplus \pi^{a_2} \oplus \cdots \oplus \pi^{a_r} \oplus {\bf 0}_{\min(m,n)-r}
  $$
  where ${\bf 0}_{\min(m,n)-r}$ is a null square matrix of size $\min(m,n)-r$.
  The matrix $S$ is called the \emph{Smith Normal Form (SNF)} of $M$, and the elements $\pi^{a_i}$ the
  \emph{invariant factors}.
\end{definition}
\begin{remark}
  The valuation of the first invariant factor of the SNF of a matrix
  $M \in \allmat_{n,n}(K)$ is the minimum of the valuations of the entries of
  $M$. In particular, the SNF of $M \in \allmat_n(\OK)$ has entries in $\OK$.
\end{remark}

@@

The \emph{associated diagonal linear programming problem} of a linear program is the iteration of the \ref{dLP} problem defined by $b' = Qb$, $c' = \left(P^{-1}\right)^T c$ and where $S$ is the Smith normal form of $A$ and $P \in GL_m(\OK)$ and $ Q \in GL_n(\OK)$ are such that $A = Q^{-1} S P$.

@@

\subsection{Definition}
%\newcommand\mat{postive semidefinite matrix } 
\newcommand\Mat{Positive semidefinite matrix }
\newcommand\mats{positive semidefinite matrices }
\newcommand\Mats{positive semidefinite matrices }

\begin{definition}
  We call a matrix $M \in \mathcal{M}_n(K)$ \emph{positive semidefinite} (in symbols $M \succeq 0$)
  if all its eigenvalues have nonnegative valuation in $\overline{K}$.
  We denote by $\mathcal{M}_n^+(K) \subset \mathcal{M}_n(K)$ the set of positive semidefinite matrices.
  The characteristic polynomial of $M$ is denoted $\chi_M$.
\end{definition}

\begin{theorem}
  \label{caracsdp}
  $M \in \mathcal{M}_n^+(K)$ if and only if $\chi_M$ has coefficients in $\OK$.
\end{theorem}
\begin{proof}
  \simone{preuve ?}
\end{proof}

\begin{corollary}\label{cor_caracsdp}
  $\mathcal{M}_n(\OK) \subset \mathcal{M}_n^+(K)$.
\end{corollary}
\begin{proof}
$\OK$ is a ring therefore the characteristic polynomial of matrix whose coefficients are $p$-adic integers have coefficients ins $Pp$ \simone{je comprends pas}. We conclude by \Cref{caracsdp}.
\end{proof}

\begin{remark}
  The reverse inclusion of \Cref{cor_caracsdp} is false. For instance the characteristic polynomial of
  $M = \left(\begin{smallmatrix} 5 + \frac{3}{5} & \frac{4}{5} \\ \frac{4}{5} & -\frac{3}{5} \end{smallmatrix}\right)$ is $\chi_M = X^2-5X-4$. However when seen as an element of $\mathbb{Q}_5$, $\chi_M \in \mathbb{Z}_5[X]$ then $M \in \mathcal{M}_2^+\left( \mathbb{Q}_5 \right)$ but no coefficient of $M$ is in $\mathbb{Z}_5 $
\end{remark}

\begin{proposition}
	The set $\mathcal{M}_n^+(K)$ is both open and closed.
\end{proposition}
\begin{proof}
Firstly, $\OK[X]$ is closed and open in $\mathbb{Q}_{p}[X]$ equiped with the infinitiy norm $\|.\|_\infty : P = \sum_{k=1}^{n} a_k X^k \to \sup |a_k|_p$. $\OK$ is indeed the open unit ball for $\|\cdot \|_\infty$ which defines a discrete distance for which open balls are also closed.
Furthermore, $\mathcal{M}^+_{n}(K) = \chi^{-1}(\OK[X]) $ where $\chi$ is the function that maps its characteristic polynomial to a matrix which is continuous.
\end{proof}

\begin{definition}
  \label{def_spectrahedra}
  Let $\mathcal{L} = A_0+\span{A_1,\ldots,A_n} \subset \mathcal{M}_n(K)$ be an affine space, for given
  matrices $A_0,A_1,\ldots,A_n$.
  The set $\mathcal{L} \cap \mathcal{M}_n^+(K)$ is called a \emph{spectrahedron} or equivalently
  its canonical preimage in $K^n$:
  $$
  S = \{x \in K^n : A_0+x_1A_1+\cdots+x_nA_n \succeq 0\}.
  $$
\end{definition}

Remark that \Cref{def_spectrahedra} is invariant under intersection: the intersection of two
(canonical preimages of) spectrahedra is a (canonical image of a) spectrahedron.

%\begin{remark}
%	As in the real case, the spectrahedron associated with the affine plan $\mathcal{L}$ will often be identified with its preimage in $K^s$ (i.e. the set of vectors $x \in K ^s$ so that the linear matrix $A(x) = A_0 + x_1A_1 + \ldots + x_sA_s$ is positive semidefinite with $A_0,A_1,\ldots,A_s$ spanning $\mathcal{L}$). 
%\end{remark}
\begin{proposition}
  A polyhedron is a spectrahedron.
\end{proposition}
\begin{proof}
  Let $\ell_i,m_j \in K[x]_1$, for $i=1,\ldots,d$ and $j=1,\ldots,e$, and let $P = Q \cap L \subset K^n$
  be the polyhedron given in \Cref{def_polyhedra}, for $Q = \{x \in K^n : \val(\ell_i) \geq 0, i=1,\ldots,d\}$
  and $L = \{x \in K^n : \val(m_j) = 0, j=1,\ldots,e\}$.
  First, denote by $A(x) = \ell_1 \oplus \cdots \oplus \ell_d$, so that $Q$ is the (preimage of the)
  spectrahedron defined by $A(x) \succeq 0$.
  Next, remark that $L \subset K^n$ is the affine space defined by equations $m_1=\ldots=m_e=0$.

  Let $\varphi_Q : K^n \to \mathcal{M}_n(K)$
\end{proof}

@@

	We call \emph{annulus} a set $A$ defined as $C = \{x \in K  | a\le \val(x) \le b\} $ for $a < b$ two nonnegative real numbers.
\end{definition}

@@

\begin{proof}
Let $A$ be the annuli defined by $a<b \in R^*_+$.

@@

Let us show that $x \in C$ if and only if  $\exists y \in K(x,y) \in \mathcal{S}$.

@@

Therefore $x \in C$ if and only if there exists $y \in K $ so that $(x, y)$ verifies \ref{eq:caracsdp}. Si $x, y$ verifies indeed \ref{caracsdp} then $\val\left(x\right)\ge a$, $\val\left(y\right)\ge -b$ and $p^{-2} \left( xy-1 \right) \ge 0$ implies that $\val\left(x\right)+\val\left(y\right)=\val\left(xy\right) = \val\left(-1\right) =0$ and therefore that $\val\left(x\right)=-\val\left(y\right)\le -(-b)  =b$, so $x \in $. Reciprocally if $x \in C$ then $(x,x^{-1}) $ verifies \ref{eq:caracsdp}  .
\end{proof}
